\subsection{Preregistration Snapshot}
\label{sec:appendix:preregistration_summary}
\label{sec:appendix:preregistration}
The preregistered analysis plan (OSF DOI: \href{https://doi.org/10.17605/OSF.IO/7D4XN}{10.17605/OSF.IO/7D4XN}) remains the authoritative record. This appendix now summarizes only the fields needed for arXiv readers; the full form, audit history, and change log stay on OSF and in the public replication repository.

\begin{itemize}[leftmargin=*]
    \item \textbf{Study question:} Does higher posting-clean L2 adoption reduce Ethereum L1 congestion? Estimand: semi-elasticity of $\log C^{fee}_t$ with respect to a 10pp adoption change over the London$\rightarrow$Merge window.
    \item \textbf{Design:} Regime-aware ITS levels + ECM with posting-clean treatment, PCA demand factor ($D^*$), regime dummies, calendar controls, and targeted shocks (mega-mints/outages). Confirmatory window: 5~Aug~2021--12~Mar~2024.
    \item \textbf{Outcomes:} Primary---median base fee; Secondary---utilization and scarcity index (report FDR-adjusted $q$ values).
    \item \textbf{Diagnostics:} Stationarity tests, Engle--Granger residual check, HAC lag justification, partial residual support plots.
    \item \textbf{Transparency:} Data and code locations (GitHub repo + Zenodo DOI) and a preregistration fidelity checklist are documented in the replication repository referenced in Appendix~\ref{sec:availability}.
\end{itemize}

Readers needing the exact prereg text should consult the OSF PDF; referencing it avoids duplicating ~15 pages and keeps the appendix budget intact.
